---
title: "Data import"
output: github_document
---

```{r setup}
library(tidyverse)
```
## Read in some data
We need to use relative paths. The benefit is that it is easier and other people can use it. Versus absolute dataset. 

Read in litters dataset. --> Fetal alcohol syndrome in pregnant mice 

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
We could all of the janitor package. But we only want the clean_names function. So we can use "package_name::". 

## Take a look at the data
Printing in the console
```{r}
litters_df
```

```{r}
head(litters_df)
tail(litters_df)
```

__Skimmer is super important to explore NA in data, also mean/sd etc__
```{r}
skimr::skim(litters_df)
```

We can use View(litters_df) in the console. 

<chr> --> this is a character variable 
<double> --> this is a number












I'm an R Markdown document! 

# Section 1

Here's a **code chunk** that samples from 
a _normal distribution_:

```{r}
samp = rnorm(100)
length(samp)
```

# Section 2

I can take the mean of the sample, too!
The mean is `r mean(samp)`.