---
title: "Data import"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```
## Read in some data
We need to use relative paths. The benefit is that it is easier and other people can use it. Versus absolute dataset. 

Read in litters dataset. --> Fetal alcohol syndrome in pregnant mice 

```{r Read csv}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
We could all of the janitor package. But we only want the clean_names function. So we can use "package_name::". 

## Take a look at the data
Printing in the console
```{r look at data}
litters_df
```

```{r Exploring data head/tail}
head(litters_df)
tail(litters_df)
```

__Skimmer is super important to explore NA in data, also mean/sd etc__
```{r skimr}
skimr::skim(litters_df)
```

We can use View(litters_df) in the console. ?read_csv

<chr> --> this is a character variable 
<double> --> this is a number

## Options to read_csv
```{r random additions to read_csv}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE, na = c("", "NA",".",999))
```

## Options to read_csv
check out '?  read_csv()' to find out more information

## Other file formats 
Read in an excel file. 

We are loading readxl. 
```{r Read Excel file}
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df
```
We can use "?read_excel". We can specify sheet, and range to import. And which cell values, using range. 

We are going to use the range function
```{r range function in read_excel}
mlb_df_2 = read_excel("./data/mlb11.xlsx", range= "A1:F7")
```

Read in SAS file 
```{r Read SAS file}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```

## Comparison with Base R

What about read.csv is built into R, but NEVER use this! Differences are relatively small but bad stuff might happen!

## Exporting data

Export the mlb sub-table 
```{r read_csv again}
write_csv(mlb_df, "./data/mlb_subtable.csv")
```

#self-assessments 
## we want to load the csv file using relative and direct manners 
```{r assessment}
FAS_pups_df = read_csv("./data/FAS_pups.csv")

FAS_pups_df = read_csv("~/Desktop/data_wrangling/data/FAS_pups.csv")
```
## Now we want to make sure column names and format are reasonable 

```{r column names and format}
FAS_pups_df = read_csv("./data/FAS_pups.csv", col_types="ciiiii") 
FAS_pups_df
skimr::skim(FAS_pups_df)

```

## Importing data using read.csv from R which is not from readr package from tidyr

```{r}
FAS_pups_df = read.csv("./data/FAS_pups.csv")
FAS_pups_df
```
We can see that all the data is displayed instead of just the first couple of rows! Which is clumsy. 



















