---
title: "Tidying Data"
output: github_document
---

#Tidying Data

## Load relevant packages & limit the lines printed in a tibble 
```{r tidyverse}
library(tidyverse)

options(tibble.print_min = 5)
```

## Create Dataframe for Public Pulse Data

```{r pulse_df dataset}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()

pulse_df
```

## Use Pivot Longer to address problem
Problem: BDI score is spread across four columns, which correspond to four observation times


```{r Pivot_longer 1}
pulse_tidy_data = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    values_to = "bdi")

pulse_tidy_data
```

We want to remove names prefix
```{r Pivot_longer 2}
pulse_tidy_data = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi")

pulse_tidy_data
```

There are also some additional transformations needed to wrap up the data wrangling process:
* like changing bl to 00m for consistency across visits 
* converting visit to a factor variable. * (It’s possible that you would want visit to be a numeric variable instead, which could be done with a different call to mutate.) 
* Lastly, it’s nice to organize the data into a reasonable order.

```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  relocate(visit) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)) %>%
  arrange(id, visit)

print(pulse_df, n = 12)
```

## Pivot_wider

Pivot_wider is the inverse of Pivot_longer

```{r create dataframe analysis_result}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

analysis_result
```
Use Pivot_wider to manage this issue!
```{r Pivot_wider}
pivot_wider(
  analysis_result, 
  names_from = "time", 
  values_from = "mean")
```

## Binding rows 
```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
```

```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>%
  pivot_longer(
    female:male,
    names_to = "gender", 
    values_to = "words") %>%
  mutate(race = str_to_lower(race)) %>% 
  select(movie, everything()) 

lotr_tidy
```

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") %>%
  janitor::clean_names() %>%
  mutate(
    sex = recode(sex, `1` = "male", `2` = "female"),
    sex = factor(sex)) 
```

```{r}
litter_data = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  relocate(litter_number) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    dose = str_to_lower(dose))

fas_data = 
  left_join(pup_data, litter_data, by = "litter_number")
```













